---
output: html_document
editor_options: 
  chunk_output_type: console
chunk_output_type: console
---





#### parameter plot
```{r}
library(tidyverse)
# fit = readRDS( "out/beta_move_0_4Dec2018.rds" )
fit = readRDS( "out/fit_20190123T195146.rds" )
pars = tibble( p1 = fit$param[,2], p2 = fit$param[,1] )
# pars <- unique( pars )
p = ggplot( pars, aes( p1, p2 ) ) + 
   stat_density_2d( aes( fill = ..level..), geom = "polygon") +
   scale_fill_continuous( low="lavenderblush", high="red" ) +
   geom_point( alpha=0.1, size=2 ) +
   # geom_jitter( alpha=0.5, size = 2) +
   # customize labels, including legend label for fill
  labs(y = "Average daily new infections of a low-infectivity case", 
       x = "Average delay before moving to another hospital (day)" ) +
  lims(x = c(0,10), y = c(0,1)) + 
  theme_classic() + theme( panel.grid.major = element_line( colour="#f0f0f0" ), legend.position = "none" ) 

library(ggExtra)
p2 = ggMarginal( p, type="histogram" )
# ggsave( "figs/delay_beta_fit.png", plot=p2, width=3.4*2, height=2.7*2, units="in" )
p2
```

### Runs across different random seeds and paramter values are combined
```{r}
# generate a data frame 
res_ctrl <- readRDS( "out/ctrl_thresholdcase_cov_20190128T122443.rds" )
pars_ctrl <- readRDS( "out/ctrl_pars_20190128T122443.rds" )
res <- readRDS( "out/thresholdcase_cov_20190128T120645.rds" )
pars <- readRDS( "out/pars_20190128T120645.rds" )
n_per_par <- pars$n_per_par[1] 
  
n <- nrow( pars_ctrl )
ci_ctrl <- as.vector( sapply( seq_len( n ), function(x) res_ctrl[[x]]$cumul_symptom_onset ) )
dur_ctrl <- as.vector( sapply( seq_len( n ), function(x) res_ctrl[[x]]$dur_outbreak ) )
num_hosp_ctrl <- as.vector( sapply( seq_len( n ), function(x) res_ctrl[[x]]$num_hospital_infected ) )

ci <- as.vector( sapply( 1:nrow(pars), function(x) res[[ x ]]$cumul_symptom_onset ) )
dur <- as.vector( sapply( 1:nrow(pars), function(x) res[[ x ]]$dur_outbreak ) )
num_hosp <- as.vector( sapply( 1:nrow(pars), function(x) res[[ x ]]$num_hospital_infected ) )
vacc <- as.vector( sapply( 1:nrow(pars), function(x)  res[[ x ]]$cumul_vacc_dose ) )  
pop_risk <- 352663
prop_vacc <- vacc / pop_risk


len <- 25
ci_ctrl <- ci_ctrl[ rep( 1:length( ci_ctrl ), each=len ) ]
dur_ctrl <- dur_ctrl[ rep( 1:length( dur_ctrl ), each=len ) ]
num_hosp_ctrl <- num_hosp_ctrl[ rep( 1:length( num_hosp_ctrl ), each=len ) ]


ve_size <- sapply( 1:nrow(pars), function(x) 100 * ( 1 - ci[ x ] / ci_ctrl[ x ] ) )
ve_dur <- sapply( 1:nrow(pars), function(x) 100 * ( 1 - dur[ x ] / dur_ctrl[ x ] ) )
ve_case_avt <- sapply( 1:nrow(pars), function(x) 10000 * ( ( ci_ctrl[ x ] - ci[ x ]  )  / vacc[ x ] ) ) #
ve_num_hosp <- sapply( 1:nrow(pars), function(x) 100 * ( 1 - num_hosp[ x ] / num_hosp_ctrl[ x ] ) )

df <- data.frame( ci=ci, dur = dur, vacc = vacc, prop_vacc = prop_vacc, 
                  num_hosp = num_hosp, ve_size = ve_size, ve_dur = ve_dur, 
                  ve_case_avt = ve_case_avt,
                  ve_num_hosp = ve_num_hosp )

pars <- pars[ rep( 1:nrow( pars ), each=pars$n_per_par[1] ), ]
resdf <-  cbind( pars, df )

library(tidyverse)
library(viridis)
df <- group_by( resdf, vacc_cov, threshold_case_vacc ) %>% 
  summarise( med_ve_size =  median( ve_size ), 
             sd_ve_size =  sd( ve_size ) )
  
ggplot( df, aes( vacc_cov, threshold_case_vacc ) ) +
  geom_tile( aes( fill = med_ve_size ) ) +
  scale_fill_viridis() + 
  annotate( "text", x=df$vacc_cov, y=df$threshold_case_vacc, label = sprintf("%.1f",df$med_ve_size) ) +
  labs( x="Vaccine coverage", y="Threshold case incidence", fill="VE" ) +
  theme_classic() 

ggplot( df, aes( prop_vacc, threshold_case_vacc ) ) +
   geom_point( aes( color = med_ve_size ) ) +
   scale_fill_viridis() +
   labs( x="Vaccine coverage", y="Threshold case incidence", fill="VE" ) +
   theme_classic() 
# 
# resdf <- resdf[ resdf$rel_pep==0.5, ]
# ## plot variance of medican across parameter estimates
# library(tidyverse)
# ggplot( resdf ) +
#   geom_boxplot( aes(as.factor(100*vacc_cov), ci, fill=as.factor(100*rel_pop) ),
#                 position = position_dodge(width = .6) ) +
#   coord_cartesian( ylim=c(0,500) ) +
#   labs( x="Vaccine coverage rate (%)", y="Cumulative incidence", 
#          fill = "Post-exposure\nprotection (%)" ) +
#   theme_classic() +
#   theme( panel.grid.major = element_line( colour="#f0f0f0" ) ) 
# # ggsave( "figs/vacccov_vacceff_inc.png", width=3.4*1.5, height=2.5*1.5, units="in" )
# 
# ggplot( resdf ) +
#   geom_boxplot( aes(as.factor(100*vacc_cov), ove, fill=as.factor(100*rel_pep) ),
#                 position = position_dodge(width = .6) ) +
#   coord_cartesian( ylim=c(-50,80) ) +
#   # geom_hline( yintercept = 70, linetype="dashed" ) +
#   labs( x="Vaccine coverage rate (%)", y="Reduction in outbreak size (%)", 
#          fill = "Post-exposure\nprotection (%)" ) +
#   theme_classic() +
#   theme( panel.grid.major = element_line( colour="#f0f0f0" ),
#          panel.grid.minor = element_line( colour="#f0f0f0" )) 
#   # annotate( "text", x=4.8, y=76, label = "Vaccine efficacy", size=3)
#    
# # ggsave( "figs/vacccov_outbreaksize_relpep.png", width=3.4*1.5, height=2.5*1.5, units="in" )
# 
# ggplot( resdf ) +
#   geom_boxplot( aes(as.factor(100*vacc_cov), dur, fill=as.factor(100*rel_pep) ),
#                 position = position_dodge(width = .6) ) +
#   # coord_cartesian( ylim=c(0,500) ) +
#   labs( x="Vaccine coverage rate (%)", y="Outbreak duration", 
#          fill = "Post-exposure\nprotection (%)" ) +
#   theme_classic() +
#   theme( panel.grid.major = element_line( colour="#f0f0f0" ) ) 
# # ggsave( "figs/vacccov_dur_relpep.png", width=3.4*1.5, height=2.5*1.5, units="in" )
# 
# ggplot( resdf ) +
#   geom_boxplot( aes(as.factor(100*vacc_cov), ove_dur, fill=as.factor(100*rel_pep) ),
#                 position = position_dodge(width = .6) ) +
#   # coord_cartesian( ylim=c(-50,80) ) +
#   # geom_hline( yintercept = 70, linetype="dashed" ) +
#   labs( x="Vaccine coverage rate (%)", y="Reduction in outbreak duration (%)", 
#          fill = "Post-exposure\nprotection (%)" ) +
#   theme_classic() +
#   theme( panel.grid.major = element_line( colour="#f0f0f0" ),
#          panel.grid.minor = element_line( colour="#f0f0f0" )) 
# 
# # ggsave( "figs/vacccov_dur_relpep.png", width=3.4*1.5, height=2.5*1.5, units="in" )
# 
# ggplot( resdf ) +
#    geom_boxplot( aes(as.factor(100*vacc_cov), case_averted, 
#                      fill=as.factor(100*rel_pep) ),
#                  position = position_dodge(width = .6) ) + 
#    coord_cartesian(ylim=c(-30,40)) +
#    labs( x="Vaccine coverage rate (%)", y="Case averted per 10,000 doses",
#          fill = "Post-exposure\nprotection (%)" ) +
#    theme_classic() +
#    theme( panel.grid.major = element_line( colour="#f0f0f0" ),
#           panel.grid.minor = element_line( colour="#f0f0f0" )) 
# 
# # ggsave( "figs/vacccov_caseaverted_relpep.png", width=3.4*1.5, height=2.5*1.5, units="in" )
# 
# ggplot( resdf ) +
#   geom_boxplot( aes(as.factor(100*vacc_cov), 
#                      vacc, fill=as.factor(100*rel_pep) ), 
#                  position = position_dodge(width = .6) ) +
#   labs( x="Vaccine coverage rate (%)", y="Number of Vaccine doses used", 
#          fill = "Vaccine\nefficacy (%)" ) +
#   theme_classic() +
#   theme( panel.grid.major = element_line( colour="#f0f0f0" ),
#          panel.grid.minor = element_line( colour="#f0f0f0" ))
#    
# # ggsave( "figs/vacccov_vaccdose_vacceff.png", width=3.4*1.5, height=2.5*1.5, units="in" )
# 
# ## 
# ggplot( resdf ) +
#   geom_boxplot( aes(as.factor(100*vacc_cov), 
#                      num_hospital, fill=as.factor(100*vacc_eff) ), 
#                  position = position_dodge(width = .6) ) +
#   labs( x="Vaccine coverage rate (%)", y="Number of affected hospitals", 
#          fill = "Vaccine\nefficacy (%)" ) +
#   theme_classic() +
  theme( panel.grid.major = element_line( colour="#f0f0f0" ) )
   
# ggsave( "figs/vacccov_vacceff_vaccdose.png", width=3.4*1.5, height=2.5*1.5, units="in" )
```
### Statistics by individuals runs for a single paramter (beta and delay_move)
```{r}
## parameters explored
# vacc_cov <- c( 0.0, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9 )
# rel_pep <- c( 0.1, 0.3, 0.5 )
# vacc_eff <- c( 0.3, 0.5, 0.7, 0.9 )
# delay_vacc <- c( 7, 14, 21, 28 )
# n_per_par <- 100
# pars <- data.frame( expand.grid( vacc_cov = vacc_cov, rel_pep = rel_pep, vacc_eff = vacc_eff, delay_vacc = delay_vacc ) )
res <- readRDS( "out/vaccscenario_thresholdcase_20181207T164825.rds" )
# saveRDS( res_cls, paste0( "out/res_control_",tstamp, ".rds" ) )
### Overall vaccine effectiveness
library(tidyverse)
## separate run required for contorl (no vaccination) -> res_Cls[[1]]
ci_red <- sapply( 1:nrow(pars), function(x) 100 * ( 1 - res[[x]]$cumul_symptom_onset / res_cls[[1]]$cumul_symptom_onset ) )
case_averted_per_10000 <- sapply( 1:nrow(pars), function(x) 10000 * ( ( res_cls[[1]]$cumul_symptom_onset - res[[x]]$cumul_symptom_onset ) / res[[x]]$cumul_vacc_dose ) )
pars_long <- pars %>% slice( rep(1:n(), each = n_per_par ) ) ## repeathe each parameter with 
resdf <- cbind( pars_long, ci_red = as.vector(ci_red), case_averted_per_10000 = as.vector(case_averted_per_10000) )

ggplot( resdf ) +
   geom_boxplot( aes(as.factor(threshold_case_vacc), ci_red, fill=as.factor(vacc_scenario) ),
                 position = position_dodge(width = .6) ) + 
   coord_cartesian(ylim=c(-100,100))+
   labs( x="Vaccine trigger incidence", y="Reduction in outbreak size (%)") +
   theme_classic() +
   theme( panel.grid.major = element_line( colour="#f0f0f0" ) )

ggsave( "figs/vacc_cov_red_ind.png", width=3.4*1.5, height=2.5*1.5, units="in" )

ggplot( resdf ) +
   geom_boxplot( aes(as.factor(threshold_case_vacc), case_averted_per_10000, fill=as.factor(vacc_scenario) ),
                 position = position_dodge(width = .6) ) + 
   coord_cartesian(ylim=c(-50,50))+
   labs( x="Vaccine trigger incidence", y="Case averted per 10,000 doses") +
   theme_classic() +
   theme( panel.grid.major = element_line( colour="#f0f0f0" ) )

ggsave( "figs/vacc_cov_case_ind.png", width=3.4*1.5, height=2.5*1.5, units="in" )

```

#### Vaccince coverage and relative efficacy of post-exposure protection 
```{r}
## 30 parametrer sets and 30 runs per parameter set 
res <- readRDS( "out/vacccov_relpop_20181218T214908.rds" )
pars <- readRDS( "out/pars_20181218T214908.rds" )

cycle <- 6 # vaccine coverage repeats every cycleth row
index_Vacc_zero <- seq( 1, nrow(pars), by = cycle ) # 6 comes from 0,0,1,0.3,...0,9
ci_vacc_zero_list <- lapply( index_Vacc_zero, function(x) matrix( rep(res[[x]]$cumul_symptom_onset, cycle), nrow=length(res[[x]]$cumul_symptom_onset) ) )
ci_control <- ci_vacc_zero_list[[1]]; 
for( i in 2:length(ci_vacc_zero_list) ) {
  ci_control <- cbind( ci_control, ci_vacc_zero_list[[i]] )
}
# overall vaccine effectiveness
# vacc_occurred <- sapply( 1:nrow(pars), function(x) (sum(res[[x]]$vacc_day_adjusted)- pars$n_per_par[1]) )
ci <- sapply( 1:nrow(pars), function(x)  res[[ x ]]$cumul_symptom_onset )
vacc <- sapply( 1:nrow(pars), function(x) res[[ x ]]$cumul_vacc_dose )  
ove <- 100 * ( 1 - ci / ci_control )
case_averted <- 10000 * ( ci_control - ci )  / vacc

df <- data.frame( ci=as.vector(t(ci)), 
                  vacc = as.vector(t(vacc)),
                  ove = as.vector(t(ove)),
                  case_averted = as.vector(t(case_averted)) )
iter <- length( res[[1]]$cumul_symptom_onset ) ## iterations acorss different random seeds during simulation
pars <- do.call( rbind, replicate( iter, pars, simplify=FALSE) ) 

resdf <-  cbind( pars, df )

# resdf <- filter( resdf, threshold_case_vacc==1, dur_vacc==2, delay_vacc == 14)


# plot variance of medican across parameter estimates
library(tidyverse)
ggplot( resdf ) +
  geom_boxplot( aes(as.factor(100*vacc_cov), ci, fill=as.factor(100*vacc_eff) ),
                position = position_dodge(width = .6) ) +
  coord_cartesian( ylim=c(0,500) ) +
  labs( x="Vaccine coverage rate (%)", y="Cumulative incidence", 
         fill = "Vaccine\nefficacy (%)" ) +
  theme_classic() +
  theme( panel.grid.major = element_line( colour="#f0f0f0" ) ) 


ggplot( resdf ) +
  geom_boxplot( aes(as.factor(100*vacc_cov), ove, fill=as.factor(100*vacc_eff) ),
                position = position_dodge(width = .6) ) +
  coord_cartesian( ylim=c(-100,100) ) +
  # geom_hline( yintercept = 70, linetype="dashed" ) +
  labs( x="Vaccine coverage rate (%)", y="Reduction in outbreak size (%)", 
         fill = "Vaccine\nefficacy (%)" ) +
  theme_classic() +
  theme( panel.grid.major = element_line( colour="#f0f0f0" ) ) 
  # annotate( "text", x=4.8, y=76, label = "Vaccine efficacy", size=3)
   
# ggsave( "figs/vacccov_vacceff.png", width=3.4*1.5, height=2.5*1.5, units="in" )

ggplot( resdf ) +
   geom_boxplot( aes(as.factor(100*vacc_cov), case_averted, 
                     fill=as.factor(100*vacc_eff) ),
                 position = position_dodge(width = .6) ) + 
   coord_cartesian(ylim=c(-30,40)) +
   labs( x="Vaccine coverage rate (%)", y="Case averted per 10,000 doses",
         fill = "Vaccine\nefficacy (%)" ) +
   theme_classic() +
   theme( panel.grid.major = element_line( colour="#f0f0f0" ) ) 

# ggsave( "figs/vacccov_caseaverted.png", width=3.4*1.5, height=2.5*1.5, units="in" )

ggplot( resdf ) +
  geom_boxplot( aes(as.factor(100*vacc_cov), 
                     vacc, fill=as.factor(100*vacc_eff) ), 
                 position = position_dodge(width = .6) ) +
  labs( x="Vaccine coverage rate (%)", y="Number of Vaccine doses used", 
         fill = "Vaccine\nefficacy (%)" ) +
  theme_classic() +
  theme( panel.grid.major = element_line( colour="#f0f0f0" ) )
   
# ggsave( "figs/vacccov_vaccdose.png", width=3.4*1.5, height=2.5*1.5, units="in" )

```

### Statistics by median or mean of individuals runs acrosss estimated beta and delay_move
#### Vaccince coverage and relative efficacy of post-exposure protection 
```{r}
res <- readRDS( "out/vacccov_relpop_par1-20_20181226T174800.rds" )
pars <- readRDS( "out/pars_par1-20_20181226T174800.rds" )
# pars <- do.call( rbind, list(pars1, pars2, pars3) )
# res <- do.call( c, list(res1, res2, res3) )
# saveRDS( res, "out/vacccov_relpop_par1-20_20181226T174800.rds" )

cycle <- 6 # 6 comes from the vaccine coverage being 0,0,1,0.3,...0,9
index_Vacc_zero <- seq( 1, nrow(pars), by=cycle ) 
ci_control <- sapply( index_Vacc_zero, function(x) median( res[[x]]$cumul_symptom_onset ) )
ci_control <- ci_control[ rep( 1:length( ci_control ), each=cycle ) ]
dur_control <- sapply( index_Vacc_zero, function(x) median( res[[x]]$dur_outbreak ) )
dur_control <- dur_control[ rep( 1:length( dur_control ), each=cycle ) ]
num_hospital_control <- sapply( index_Vacc_zero, function(x) median( res[[x]]$num_hospital_infected ) )
num_hospital_control <- num_hospital_control[ rep( 1:length( num_hospital_control ), each=cycle ) ]

ci <- sapply( 1:nrow(pars), function(x) median( res[[ x ]]$cumul_symptom_onset) )
dur <- sapply( 1:nrow(pars), function(x) median( res[[ x ]]$dur_outbreak ) )
num_hospital <- sapply( 1:nrow(pars), function(x) median( res[[ x ]]$num_hospital_infected ) )
vacc <- sapply( 1:nrow(pars), function(x) median( res[[ x ]]$cumul_vacc_dose ) )  
ove <- sapply( 1:nrow(pars), function(x) 100 * ( 1 - ci[ x ] / ci_control[ x ] ) )
ove_dur <- sapply( 1:nrow(pars), function(x) 100 * ( 1 - dur[ x ] / dur_control[ x ] ) )

case_averted <- sapply( 1:nrow(pars), function(x) 10000 * ( ( ci_control[ x ] - ci[ x ]  )  / vacc[ x ] ) ) #
ove_hospital <- sapply( 1:nrow(pars), function(x) 100 * ( 1 - num_hospital[ x ] / num_hospital_control[ x ] ) )

df <- data.frame( ci=ci, vacc = vacc, ove = ove, dur = dur, ove_dur = ove_dur, 
                  case_averted = case_averted,
                  num_hospital = num_hospital, ove_hospital = ove_hospital )

resdf <-  cbind( pars, df )

resdf <- resdf[ resdf$rel_pep==0.5, ]
## plot variance of medican across parameter estimates
library(tidyverse)
ggplot( resdf ) +
  geom_boxplot( aes(as.factor(100*vacc_cov), ci, fill=as.factor(100*rel_pop) ),
                position = position_dodge(width = .6) ) +
  coord_cartesian( ylim=c(0,500) ) +
  labs( x="Vaccine coverage rate (%)", y="Cumulative incidence", 
         fill = "Post-exposure\nprotection (%)" ) +
  theme_classic() +
  theme( panel.grid.major = element_line( colour="#f0f0f0" ) ) 
# ggsave( "figs/vacccov_vacceff_inc.png", width=3.4*1.5, height=2.5*1.5, units="in" )

ggplot( resdf ) +
  geom_boxplot( aes(as.factor(100*vacc_cov), ove, fill=as.factor(100*rel_pep) ),
                position = position_dodge(width = .6) ) +
  coord_cartesian( ylim=c(-50,80) ) +
  # geom_hline( yintercept = 70, linetype="dashed" ) +
  labs( x="Vaccine coverage rate (%)", y="Reduction in outbreak size (%)", 
         fill = "Post-exposure\nprotection (%)" ) +
  theme_classic() +
  theme( panel.grid.major = element_line( colour="#f0f0f0" ),
         panel.grid.minor = element_line( colour="#f0f0f0" )) 
  # annotate( "text", x=4.8, y=76, label = "Vaccine efficacy", size=3)
   
# ggsave( "figs/vacccov_outbreaksize_relpep.png", width=3.4*1.5, height=2.5*1.5, units="in" )

ggplot( resdf ) +
  geom_boxplot( aes(as.factor(100*vacc_cov), dur, fill=as.factor(100*rel_pep) ),
                position = position_dodge(width = .6) ) +
  # coord_cartesian( ylim=c(0,500) ) +
  labs( x="Vaccine coverage rate (%)", y="Outbreak duration", 
         fill = "Post-exposure\nprotection (%)" ) +
  theme_classic() +
  theme( panel.grid.major = element_line( colour="#f0f0f0" ) ) 
# ggsave( "figs/vacccov_dur_relpep.png", width=3.4*1.5, height=2.5*1.5, units="in" )

ggplot( resdf ) +
  geom_boxplot( aes(as.factor(100*vacc_cov), ove_dur, fill=as.factor(100*rel_pep) ),
                position = position_dodge(width = .6) ) +
  # coord_cartesian( ylim=c(-50,80) ) +
  # geom_hline( yintercept = 70, linetype="dashed" ) +
  labs( x="Vaccine coverage rate (%)", y="Reduction in outbreak duration (%)", 
         fill = "Post-exposure\nprotection (%)" ) +
  theme_classic() +
  theme( panel.grid.major = element_line( colour="#f0f0f0" ),
         panel.grid.minor = element_line( colour="#f0f0f0" )) 

# ggsave( "figs/vacccov_dur_relpep.png", width=3.4*1.5, height=2.5*1.5, units="in" )

ggplot( resdf ) +
   geom_boxplot( aes(as.factor(100*vacc_cov), case_averted, 
                     fill=as.factor(100*rel_pep) ),
                 position = position_dodge(width = .6) ) + 
   coord_cartesian(ylim=c(-30,40)) +
   labs( x="Vaccine coverage rate (%)", y="Case averted per 10,000 doses",
         fill = "Post-exposure\nprotection (%)" ) +
   theme_classic() +
   theme( panel.grid.major = element_line( colour="#f0f0f0" ),
          panel.grid.minor = element_line( colour="#f0f0f0" )) 

# ggsave( "figs/vacccov_caseaverted_relpep.png", width=3.4*1.5, height=2.5*1.5, units="in" )

ggplot( resdf ) +
  geom_boxplot( aes(as.factor(100*vacc_cov), 
                     vacc, fill=as.factor(100*rel_pep) ), 
                 position = position_dodge(width = .6) ) +
  labs( x="Vaccine coverage rate (%)", y="Number of Vaccine doses used", 
         fill = "Vaccine\nefficacy (%)" ) +
  theme_classic() +
  theme( panel.grid.major = element_line( colour="#f0f0f0" ),
         panel.grid.minor = element_line( colour="#f0f0f0" ))
   
# ggsave( "figs/vacccov_vaccdose_vacceff.png", width=3.4*1.5, height=2.5*1.5, units="in" )

## 
ggplot( resdf ) +
  geom_boxplot( aes(as.factor(100*vacc_cov), 
                     num_hospital, fill=as.factor(100*vacc_eff) ), 
                 position = position_dodge(width = .6) ) +
  labs( x="Vaccine coverage rate (%)", y="Number of affected hospitals", 
         fill = "Vaccine\nefficacy (%)" ) +
  theme_classic() +
  theme( panel.grid.major = element_line( colour="#f0f0f0" ) )
   
# ggsave( "figs/vacccov_vacceff_vaccdose.png", width=3.4*1.5, height=2.5*1.5, units="in" )
```






#### Vaccince coverage and isolation rate 
```{r}
## 30 parametrer sets and 200 runs per parameter set 
res1 <- readRDS( "out/isolation_vacccov_20181225T202552_par1.rds" )
res2 <- readRDS( "out/isolation_vacccov_20181225T214951_par2-6.rds" )
pars1 <- readRDS( "out/pars_20181225T202552.rds" )
pars2 <- readRDS( "out/pars_20181225T214951.rds" )

pars <- do.call( rbind, list(pars1, pars2) )
res <- do.call( c, list(res1, res2) )

pars_2400 <- pars[ seq(1,72,2), ]
pars_4259 <- pars[ seq(2,72,2), ]
pars <- do.call( rbind, list(pars_2400, pars_4259) )
res_2400 <- lapply( seq(1,72,2), function(x) res[[x]] )
res_4259 <- lapply( seq(2,72,2), function(x) res[[x]] )
res <- do.call( c, list(res_2400, res_4259) )

cycle <- 6 # 6 comes from the vaccine coverage being 0,0,1,0.3,...0,9
index_Vacc_zero <- seq( 1, nrow(pars), by=cycle ) 
ci_control <- sapply( index_Vacc_zero, function(x) median( res[[x]]$cumul_symptom_onset ) )
ci_control <- ci_control[ rep( 1:length( ci_control ), each=cycle ) ]
dur_control <- sapply( index_Vacc_zero, function(x) median( res[[x]]$dur_outbreak ) )
dur_control <- dur_control[ rep( 1:length( dur_control ), each=cycle ) ]
num_hospital_control <- sapply( index_Vacc_zero, function(x) median( res[[x]]$num_hospital_infected ) )
num_hospital_control <- num_hospital_control[ rep( 1:length( num_hospital_control ), each=cycle ) ]

ci <- sapply( 1:nrow(pars), function(x) median( res[[ x ]]$cumul_symptom_onset) )
dur <- sapply( 1:nrow(pars), function(x) median( res[[ x ]]$dur_outbreak ) )
num_hospital <- sapply( 1:nrow(pars), function(x) median( res[[ x ]]$num_hospital_infected ) )
vacc <- sapply( 1:nrow(pars), function(x) median( res[[ x ]]$cumul_vacc_dose ) )  
ove <- sapply( 1:nrow(pars), function(x) 100 * ( 1 - ci[ x ] / ci_control[ x ] ) )
ove_dur <- sapply( 1:nrow(pars), function(x) 100 * ( 1 - dur[ x ] / dur_control[ x ] ) )
case_averted <- sapply( 1:nrow(pars), function(x) 10000 * ( ( ci_control[ x ] - ci[ x ]  )  / vacc[ x ] ) ) #
ove_hospital <- sapply( 1:nrow(pars), function(x) 100 * ( 1 - num_hospital[ x ] / num_hospital_control[ x ] ) )

df <- data.frame( ci=ci, vacc = vacc, ove = ove, dur = dur, ove_dur = ove_dur, 
                  case_averted = case_averted,
                  num_hospital = num_hospital, ove_hospital = ove_hospital )

resdf <-  cbind( pars, df )

# resdf <- resdf[ resdf$vacc_scenario=="Distance", ]
## plot variance of medican across parameter estimates
library(tidyverse)
ggplot( resdf ) +
  geom_boxplot( aes(as.factor(100*vacc_cov), ci, fill=as.factor(meantime_isolation_middle) ),
                position = position_dodge(width = .6) ) +
  coord_cartesian( ylim=c(0,750) ) +
  labs( x="Vaccine coverage rate (%)", y="Cumulative incidence", 
         fill = "Isolation rate" ) +
    scale_fill_manual(labels = c("Baseline", "Slow"), values = c("#F8766D", "#00BFC4")) + 
  theme_classic() +
  theme( panel.grid.major = element_line( colour="#f0f0f0" ) ) 
# ggsave( "figs/vacccov_inc_isolation.png", width=3.4*1.5, height=2.5*1.5, units="in" )

ggplot( resdf ) +
  geom_boxplot( aes(as.factor(100*vacc_cov), ove, fill=as.factor(meantime_isolation_middle) ),
                position = position_dodge(width = .6) ) +
  # coord_cartesian( ylim=c(-50,80) ) +
  # geom_hline( yintercept = 70, linetype="dashed" ) +
  labs( x="Vaccine coverage rate (%)", y="Reduction in outbreak size (%)", 
         fill = "Isolation rate" ) +
  scale_fill_manual(labels = c("Baseline", "Slow"), values = c("#F8766D", "#00BFC4")) + 
  theme_classic() +
  theme( panel.grid.major = element_line( colour="#f0f0f0" ),
         panel.grid.minor = element_line( colour="#f0f0f0" )) 
  # annotate( "text", x=4.8, y=76, label = "Vaccine efficacy", size=3)
   
# ggsave( "figs/vacccov_outbreaksize_isolationrate.png", width=3.4*1.5, height=2.5*1.5, units="in" )

ggplot( resdf ) +
  geom_boxplot( aes(as.factor(100*vacc_cov), dur, fill=as.factor(meantime_isolation_middle) ),
                position = position_dodge(width = .6) ) +
  # coord_cartesian( ylim=c(0,500) ) +
  labs( x="Vaccine coverage rate (%)", y="Outbreak duration", 
         fill = "Isolation rate" ) +
  scale_fill_manual(labels = c("Baseline", "Slow"), values = c("#F8766D", "#00BFC4")) + 
  theme_classic() +
  theme( panel.grid.major = element_line( colour="#f0f0f0" ) ) 
# ggsave( "figs/vacccov_dur_relpep.png", width=3.4*1.5, height=2.5*1.5, units="in" )


ggplot( resdf ) +
  geom_boxplot( aes(as.factor(100*vacc_cov), ove_dur, fill=as.factor(meantime_isolation_middle) ),
                position = position_dodge(width = .6) ) +
  # coord_cartesian( ylim=c(-50,80) ) +
  # geom_hline( yintercept = 70, linetype="dashed" ) +
  scale_fill_manual(labels = c("Baseline", "Slow"), values = c("#F8766D", "#00BFC4")) + 
  labs( x="Vaccine coverage rate (%)", y="Reduction in outbreak duration (%)", 
         fill = "Isolation rate" ) +
  theme_classic() +
  theme( panel.grid.major = element_line( colour="#f0f0f0" ),
         panel.grid.minor = element_line( colour="#f0f0f0" )) 

# ggsave( "figs/vacccov_dur_isolation.png", width=3.4*1.5, height=2.5*1.5, units="in" )

ggplot( resdf ) +
   geom_boxplot( aes(as.factor(100*vacc_cov), case_averted, 
                     fill=as.factor(meantime_isolation_middle) ),
                 position = position_dodge(width = .6) ) + 
   # coord_cartesian(ylim=c(-30,40)) +
   labs( x="Vaccine coverage rate (%)", y="Case averted per 10,000 doses",
         fill = "Isolation rate" ) +
  scale_fill_manual(labels = c("Baseline", "Slow"), values = c("#F8766D", "#00BFC4")) +
   theme_classic() +
   theme( panel.grid.major = element_line( colour="#f0f0f0" ),
          panel.grid.minor = element_line( colour="#f0f0f0" )) 

# ggsave( "figs/vacccov_caseaverted_isolation.png", width=3.4*1.5, height=2.5*1.5, units="in" )

ggplot( resdf ) +
  geom_boxplot( aes(as.factor(100*vacc_cov), 
                     vacc, fill=as.factor(meantime_isolation_middle) ), 
                 position = position_dodge(width = .6) ) +
  labs( x="Vaccine coverage rate (%)", y="Number of Vaccine doses used", 
         fill = "Vaccine\nefficacy (%)" ) +
  theme_classic() +
  theme( panel.grid.major = element_line( colour="#f0f0f0" ),
         panel.grid.minor = element_line( colour="#f0f0f0" ))
   
# ggsave( "figs/vacccov_vaccdose_vacceff.png", width=3.4*1.5, height=2.5*1.5, units="in" )

## 
ggplot( resdf ) +
  geom_boxplot( aes(as.factor(100*vacc_cov), 
                     num_hospital, fill=as.factor(meantime_isolation_middle) ), 
                 position = position_dodge(width = .6) ) +
  labs( x="Vaccine coverage rate (%)", y="Number of affected hospitals", 
         fill = "Vaccine\nefficacy (%)" ) +
  theme_classic() +
  theme( panel.grid.major = element_line( colour="#f0f0f0" ) )
   
# ggsave( "figs/vacccov_vacceff_vaccdose.png", width=3.4*1.5, height=2.5*1.5, units="in" )

```


#### Vaccination scenario 
```{r}
## 30 parametrer sets and 200 runs per parameter set 
res <- readRDS( "out/vacccov_relpop_20181219T190631.rds" )
res <- readRDS( "out/vacccov_relpop_par1-20_20181226T174800.rds")
pars <- readRDS( "out/pars_par1-20_20181226T174800.rds")
pars <- readRDS( "out/pars_20181219T190631.rds" )

cycle <- 6# 6 comes from the vaccine coverage being 0,0,1,0.3,...0,9
index_Vacc_zero <- seq( 1, nrow(pars), by=cycle ) 
ci_control <- sapply( index_Vacc_zero, function(x) median( res[[x]]$cumul_symptom_onset ) )
ci_control <- ci_control[ rep( 1:length( ci_control ), each=cycle ) ]
# overall vaccine effectiveness
# vacc_occurred <- sapply( 1:nrow(pars), function(x) (sum(res[[x]]$vacc_day_adjusted)- pars$n_per_par[1]) )
ci <- sapply( 1:nrow(pars), function(x) median( res[[ x ]]$cumul_symptom_onset) )
vacc <- sapply( 1:nrow(pars), function(x) median( res[[ x ]]$cumul_vacc_dose ) )  
ove <- sapply( 1:nrow(pars), function(x) 100 * ( 1 - ci[ x ] / ci_control[ x ] ) )
case_averted <- sapply( 1:nrow(pars), function(x) 10000 * ( ( ci_control[ x ] - ci[ x ]  )  / vacc[ x ] ) ) #

df <- data.frame( ci=ci, vacc = vacc, ove = ove, case_averted = case_averted)

resdf <-  cbind( pars, df )

# resdf <- dplyr::filter( resdf, vacc_eff == 0.3 )
resdf <- resdf[ resdf$vacc_eff==0.7, ]
## plot variance of medican across parameter estimates
library(ggplot2)

ggplot( resdf ) +
  geom_boxplot( aes(as.factor(100*vacc_cov), ci, fill=as.factor(vacc_scenario) ),
                position = position_dodge(width = .6) ) +
  coord_cartesian( ylim=c(0,500) ) +
  labs( x="Vaccine coverage rate (%)", y="Cumulative incidence", 
         fill = "Vaccination\nscenario" ) +
  theme_classic() +
  theme( panel.grid.major = element_line( colour="#f0f0f0" ) ) 
ggsave( "figs/vacccov_inc_vaccscenario_ve07.png", width=3.4*1.5, height=2.5*1.5, units="in" )


ggplot( resdf ) +
  geom_boxplot( aes(as.factor(100*vacc_cov), ove, fill=as.factor(vacc_scenario) ),
                position = position_dodge(width = .6) ) +
  coord_cartesian( ylim=c(-50,80) ) +
  # geom_hline( yintercept = 70, linetype="dashed" ) +
  labs( x="Vaccine coverage rate (%)", y="Reduction in outbreak size (%)", 
         fill = "Vaccination\nscenario" ) +
  theme_classic() +
  theme( panel.grid.major = element_line( colour="#f0f0f0" ), 
         panel.grid.minor = element_line( colour="#f0f0f0" )) 
  # annotate( "text", x=4.8, y=76, label = "Vaccine efficacy", size=3)
   
ggsave( "figs/vacccov_outbreaksize_vaccscenario_ve07.png", width=3.4*1.5, height=2.5*1.5, units="in" )

ggplot( resdf ) +
   geom_boxplot( aes(as.factor(100*vacc_cov), case_averted, 
                     fill=as.factor(vacc_scenario) ),
                 position = position_dodge(width = .6) ) + 
   coord_cartesian(ylim=c(-30,70)) +
   labs( x="Vaccine coverage rate (%)", y="Case averted per 10,000 doses",
         fill = "Vaccination\nscenario" ) +
   theme_classic() +
   theme( panel.grid.major = element_line( colour="#f0f0f0" ) ) 

ggsave( "figs/vacccov_caseaverted_vaccscenario_ve07.png", width=3.4*1.5, height=2.5*1.5, units="in" )

ggplot( resdf ) +
  geom_boxplot( aes(as.factor(100*vacc_cov), 
                     vacc, fill=as.factor(vacc_scenario) ), 
                 position = position_dodge(width = .6) ) +
  labs( x="Vaccine coverage rate (%)", y="Number of Vaccine doses used", 
         fill = "Vaccination\nscenario" ) +
  theme_classic() +
  theme( panel.grid.major = element_line( colour="#f0f0f0" ) )
   
ggsave( "figs/vacccov_vaccdose_vaccscenario_ve07.png", width=3.4*1.5, height=2.5*1.5, units="in" )
```


#### Importation we need to account for whether vaccination occurred or not
```{r}
## 30 parametrer sets and 200 runs per parameter set 
res <- readRDS( "out/importation_vacccov_20181221T000028.rds" )
pars <- readRDS( "out/pars_20181221T000028.rds" )

cycle <- 6# 6 comes from the vaccine coverage being 0,0,1,0.3,...0,9
index_Vacc_zero <- seq( 1, nrow(pars), by=cycle ) 
ci_control <- sapply( index_Vacc_zero, function(x) median( res[[x]]$cumul_symptom_onset[ which(res[[x]]$vacc_day_adjusted==1) ] ) )
ci_control <- ci_control[ rep( 1:length( ci_control ), each=cycle ) ]

dur_control <- sapply( index_Vacc_zero, function(x) median( res[[x]]$dur_outbreak[ which(res[[x]]$vacc_day_adjusted==1) ] ) )
dur_control <- dur_control[ rep( 1:length( dur_control ), each=cycle ) ]

num_hospital_control <- sapply( index_Vacc_zero, function(x) median( res[[x]]$num_hospital_infected[ which(res[[x]]$vacc_day_adjusted==1) ] ) )
num_hospital_control <- num_hospital_control[ rep( 1:length( num_hospital_control ), each=cycle ) ]

# overall vaccine effectiveness
ci <- sapply( 1:nrow(pars), function(x) median( res[[ x ]]$cumul_symptom_onset[ which(res[[x]]$vacc_day_adjusted==1) ]) )
dur <- sapply( 1:nrow(pars), function(x) median( res[[ x ]]$dur_outbreak[ which(res[[x]]$vacc_day_adjusted==1) ] ) )
num_hospital <- sapply( 1:nrow(pars), function(x) median( res[[ x ]]$num_hospital_infected[ which(res[[x]]$vacc_day_adjusted==1) ] ) )
vacc <- sapply( 1:nrow(pars), function(x) median( res[[ x ]]$cumul_vacc_dose[ which(res[[x]]$vacc_day_adjusted==1) ] ) )  

ove <- sapply( 1:nrow(pars), function(x) 100 * ( 1 - ci[ x ] / ci_control[ x ] ) )
ove_dur <- sapply( 1:nrow(pars), function(x) 100 * ( 1 - dur[ x ] / dur_control[ x ] ) )
case_averted <- sapply( 1:nrow(pars), function(x) 10000 * ( ( ci_control[ x ] - ci[ x ]  )  / vacc[ x ] ) ) #
ove_hospital <- sapply( 1:nrow(pars), function(x) 100 * ( 1 - num_hospital[ x ] / num_hospital_control[ x ] ) )

df <- data.frame( ci=ci, vacc = vacc, ove = ove, case_averted = case_averted,
                  num_hospital = num_hospital, ove_hospital = ove_hospital )

resdf <-  cbind( pars, df )

resdf <- resdf[ resdf$vacc_scenario=="Distance", ]
## plot variance of medican across parameter estimates
library(tidyverse)

ggplot( resdf ) +
  geom_boxplot( aes(as.factor(100*vacc_cov), ci, fill=as.factor(100*vacc_eff) ),
                position = position_dodge(width = .6) ) +
  coord_cartesian( ylim=c(0,2000) ) +
  labs( x="Vaccine coverage rate (%)", y="Cumulative incidence", 
         fill = "Vaccine\nefficacy (%)" ) +
  theme_classic() +
  theme( panel.grid.major = element_line( colour="#f0f0f0" ) ) 
# ggsave( "figs/vacccov_vacceff_inc.png", width=3.4*1.5, height=2.5*1.5, units="in" )


ggplot( resdf ) +
  geom_boxplot( aes(as.factor(100*vacc_cov), ove, fill=as.factor(100*vacc_eff) ),
                position = position_dodge(width = .6) ) +
  coord_cartesian( ylim=c(-50,80) ) +
  # geom_hline( yintercept = 70, linetype="dashed" ) +
  labs( x="Vaccine coverage rate (%)", y="Reduction in outbreak size (%)", 
         fill = "Vaccine\nefficacy (%)" ) +
  theme_classic() +
  theme( panel.grid.major = element_line( colour="#f0f0f0" ),
         panel.grid.minor = element_line( colour="#f0f0f0" )) 
  # annotate( "text", x=4.8, y=76, label = "Vaccine efficacy", size=3)
   
ggsave( "figs/importation_vacccov_outbreaksize.png", width=3.4*1.5, height=2.5*1.5, units="in" )


ggplot( resdf ) +
  geom_boxplot( aes(as.factor(100*vacc_cov), dur, fill=as.factor(100*vacc_eff) ),
                position = position_dodge(width = .6) ) +
  # coord_cartesian( ylim=c(0,500) ) +
  labs( x="Vaccine coverage rate (%)", y="Outbreak duration", 
         fill = "Vaccine\nefficacy (%)" ) +
  theme_classic() +
  theme( panel.grid.major = element_line( colour="#f0f0f0" ) ) 
# ggsave( "figs/vacccov_vacceff_inc.png", width=3.4*1.5, height=2.5*1.5, units="in" )


ggplot( resdf ) +
  geom_boxplot( aes(as.factor(100*vacc_cov), ove_dur, fill=as.factor(100*vacc_eff) ),
                position = position_dodge(width = .6) ) +
  coord_cartesian( ylim=c(-20,20) ) +
  # geom_hline( yintercept = 70, linetype="dashed" ) +
  labs( x="Vaccine coverage rate (%)", y="Reduction in outbreak duration (%)", 
         fill = "Vaccine\nefficacy (%)" ) +
  theme_classic() +
  theme( panel.grid.major = element_line( colour="#f0f0f0" ),
         panel.grid.minor = element_line( colour="#f0f0f0" )) 

ggsave( "figs/outbreakdur_importation_vacccov.png", width=3.4*1.5, height=2.5*1.5, units="in" )

ggplot( resdf ) +
   geom_boxplot( aes(as.factor(100*vacc_cov), case_averted, 
                     fill=as.factor(100*vacc_eff) ),
                 position = position_dodge(width = .6) ) + 
   coord_cartesian(ylim=c(-30,200)) +
   labs( x="Vaccine coverage rate (%)", y="Case averted per 10,000 doses",
         fill = "Vaccine\nefficacy (%)" ) +
   theme_classic() +
   theme( panel.grid.major = element_line( colour="#f0f0f0" ),
          panel.grid.minor = element_line( colour="#f0f0f0" )) 

ggsave( "figs/caseaverted_importation_vacccov.png", width=3.4*1.5, height=2.5*1.5, units="in" )

ggplot( resdf ) +
  geom_boxplot( aes(as.factor(100*vacc_cov), 
                     vacc, fill=as.factor(100*vacc_eff) ), 
                 position = position_dodge(width = .6) ) +
  labs( x="Vaccine coverage rate (%)", y="Number of Vaccine doses used", 
         fill = "Vaccine\nefficacy (%)" ) +
  theme_classic() +
  theme( panel.grid.major = element_line( colour="#f0f0f0" ),
         panel.grid.minor = element_line( colour="#f0f0f0" ))
   
ggsave( "figs/vaccdose_importation_vacccov.png", width=3.4*1.5, height=2.5*1.5, units="in" )

## 
ggplot( resdf ) +
  geom_boxplot( aes(as.factor(100*vacc_cov), 
                     ove_hospital, fill=as.factor(100*vacc_eff) ), 
                 position = position_dodge(width = .6) ) +
  labs( x="Vaccine coverage rate (%)", y="Reduction in the number of affected hospitals (%)", 
         fill = "Vaccine\nefficacy (%)" ) +
  coord_cartesian(ylim=c(-50,50)) +
  theme_classic() +
  theme( panel.grid.major = element_line( colour="#f0f0f0" ) )
   
# ggsave( "figs/vacccov_vacceff_vaccdose.png", width=3.4*1.5, height=2.5*1.5, units="in" )

```
#### Other statistics (variance-to-mean ratio, number of hospitals affected )  
```{r}
## 30 parametrer sets and 200 runs per parameter set 
res <- readRDS( "out/vacccov_relpop_20181219T190631.rds" )
pars <- readRDS( "out/pars_20181219T190631.rds" )

cycle <- 6# 6 comes from the vaccine coverage being 0,0,1,0.3,...0,9
index_Vacc_zero <- seq( 1, nrow(pars), by=cycle ) 
ci_control <- sapply( index_Vacc_zero, function(x) median( res[[x]]$cumul_symptom_onset ) )
ci_control <- ci_control[ rep( 1:length( ci_control ), each=cycle ) ]
num_hospital_control <- sapply( index_Vacc_zero, function(x) median( res[[x]]$num_hospital_infected ) )
num_hospital_control <- num_hospital_control[ rep( 1:length( num_hospital_control ), each=cycle ) ]


# overall vaccine effectiveness
# vacc_occurred <- sapply( 1:nrow(pars), function(x) (sum(res[[x]]$vacc_day_adjusted)- pars$n_per_par[1]) )
ci <- sapply( 1:nrow(pars), function(x) median( res[[ x ]]$cumul_symptom_onset) )
num_hospital <- sapply( 1:nrow(pars), function(x) median( res[[ x ]]$num_hospital_infected ) )
vacc <- sapply( 1:nrow(pars), function(x) median( res[[ x ]]$cumul_vacc_dose ) )  
ove <- sapply( 1:nrow(pars), function(x) 100 * ( 1 - ci[ x ] / ci_control[ x ] ) )
case_averted <- sapply( 1:nrow(pars), function(x) 10000 * ( ( ci_control[ x ] - ci[ x ]  )  / vacc[ x ] ) ) #
red_hospital <- sapply( 1:nrow(pars), function(x) 100 * ( 1 - num_hospital[ x ] / num_hospital_control[ x ] ) )

df <- data.frame( ci=ci, vacc = vacc, ove = ove, case_averted = case_averted,
                  num_hospital = num_hospital, red_hospital = red_hospital )

resdf <-  cbind( pars, df )

resdf <- resdf[ resdf$vacc_scenario=="Distance", ]
## plot variance of medican across parameter estimates
library(tidyverse)

ggplot( resdf ) +
  geom_boxplot( aes(as.factor(100*vacc_cov), ci, fill=as.factor(100*vacc_eff) ),
                position = position_dodge(width = .6) ) +
  coord_cartesian( ylim=c(0,500) ) +
  labs( x="Vaccine coverage rate (%)", y="Cumulative incidence", 
         fill = "Vaccine\nefficacy (%)" ) +
  theme_classic() +
  theme( panel.grid.major = element_line( colour="#f0f0f0" ) ) 
# ggsave( "figs/vacccov_vacceff_inc.png", width=3.4*1.5, height=2.5*1.5, units="in" )


ggplot( resdf ) +
  geom_boxplot( aes(as.factor(100*vacc_cov), ove, fill=as.factor(100*vacc_eff) ),
                position = position_dodge(width = .6) ) +
  coord_cartesian( ylim=c(-50,80) ) +
  # geom_hline( yintercept = 70, linetype="dashed" ) +
  labs( x="Vaccine coverage rate (%)", y="Reduction in outbreak size (%)", 
         fill = "Vaccine\nefficacy (%)" ) +
  theme_classic() +
  theme( panel.grid.major = element_line( colour="#f0f0f0" ),
         panel.grid.minor = element_line( colour="#f0f0f0" )) 
  # annotate( "text", x=4.8, y=76, label = "Vaccine efficacy", size=3)
   
ggsave( "figs/vacccov_outbreaksize_vacceff.png", width=3.4*1.5, height=2.5*1.5, units="in" )

ggplot( resdf ) +
   geom_boxplot( aes(as.factor(100*vacc_cov), case_averted, 
                     fill=as.factor(100*vacc_eff) ),
                 position = position_dodge(width = .6) ) + 
   coord_cartesian(ylim=c(-30,40)) +
   labs( x="Vaccine coverage rate (%)", y="Case averted per 10,000 doses",
         fill = "Vaccine\nefficacy (%)" ) +
   theme_classic() +
   theme( panel.grid.major = element_line( colour="#f0f0f0" ),
          panel.grid.minor = element_line( colour="#f0f0f0" )) 

ggsave( "figs/vacccov_caseaverted_vacceff.png", width=3.4*1.5, height=2.5*1.5, units="in" )

ggplot( resdf ) +
  geom_boxplot( aes(as.factor(100*vacc_cov), 
                     vacc, fill=as.factor(100*vacc_eff) ), 
                 position = position_dodge(width = .6) ) +
  labs( x="Vaccine coverage rate (%)", y="Number of Vaccine doses used", 
         fill = "Vaccine\nefficacy (%)" ) +
  theme_classic() +
  theme( panel.grid.major = element_line( colour="#f0f0f0" ),
         panel.grid.minor = element_line( colour="#f0f0f0" ))
   
ggsave( "figs/vacccov_vaccdose_vacceff.png", width=3.4*1.5, height=2.5*1.5, units="in" )

## 
ggplot( resdf ) +
  geom_boxplot( aes(as.factor(100*vacc_cov), 
                     num_hospital, fill=as.factor(100*vacc_eff) ), 
                 position = position_dodge(width = .6) ) +
  labs( x="Vaccine coverage rate (%)", y="Number of affected hospitals", 
         fill = "Vaccine\nefficacy (%)" ) +
  theme_classic() +
  theme( panel.grid.major = element_line( colour="#f0f0f0" ) )
   
# ggsave( "figs/vacccov_vacceff_vaccdose.png", width=3.4*1.5, height=2.5*1.5, units="in" )

```

```{r}
## 30 parametrer sets and 30 runs per parameter set 
res <- readRDS( "out/vacccov_relpop_20181212T090546.rds" )
pars <- readRDS( "out/pars_20181212T090546.rds" )
# res_control <- readRDS( "out/control_201C81211T131742.rds" ) 

# beta_repeat <- 15
# index_Vacc_zero <- seq( 1, nrow(pars), by=beta_repeat )
index_Vacc_zero <- seq( 1, nrow(pars), by=6 )
# med_inc_control <- sapply( seq_along(res_control), function(x) median( res_control[[x]]$cumul_symptom_onset ) ) # zero vaccine coverage
med_inc_control <- sapply( index_Vacc_zero, function(x) median( res[[x]]$cumul_symptom_onset ) )
# beta values same over 15 sim runs
med_inc_control <- med_inc_control[ rep( 1:length( med_inc_control ), each=6 ) ]
# overall vaccine effectiveness
# vacc_occurred <- sapply( 1:nrow(pars), function(x) (sum(res[[x]]$vacc_day_adjusted)- pars$n_per_par[1]) )
ci <- sapply( 1:nrow(pars), function(x) median(res[[x]]$cumul_symptom_onset) )
vacc <- sapply( 1:nrow(pars), function(x) median( res[[x]]$cumul_vacc_dose ) )  
ove <- sapply( 1:nrow(pars), function(x) 100 * ( 1 - ci[x] / med_inc_control[ x ] ) )
case_averted <- sapply( 1:nrow(pars), function(x) 10000 * ( ( med_inc_control[ x ] - ci[ x ]  )  / vacc[ x ] ) ) #

df <- data.frame( ci=ci, vacc = vacc, ove = ove, case_averted = case_averted)
resdf <-  cbind( pars, df )

## plot variance of medican across parameter estimates
library( ggplot2 )
ggplot( resdf ) +
  geom_boxplot( aes(as.factor(100*vacc_cov), 
                     ove, fill=as.factor(100*vacc_eff) ), 
                 position = position_dodge(width = .6) ) +
  coord_cartesian( ylim=c(-100,100) ) +
  # geom_hline( yintercept = 70, linetype="dashed" ) +
  labs( x="Vaccine coverage rate (%)", y="Reduction in outbreak size (%)", 
         fill = "Vaccine\nefficacy (%)" ) +
  theme_classic() +
  theme( panel.grid.major = element_line( colour="#f0f0f0" ) ) 
  # annotate( "text", x=4.8, y=76, label = "Vaccine efficacy", size=3)
   
ggsave( "figs/vacccov_vacceff.png", width=3.4*1.5, height=2.5*1.5, units="in" )

ggplot( resdf ) +
   geom_boxplot( aes(as.factor(100*vacc_cov), case_averted, 
                     fill=as.factor(100*vacc_eff) ),
                 position = position_dodge(width = .6) ) + 
   coord_cartesian(ylim=c(-30,40)) +
   labs( x="Vaccine coverage rate (%)", y="Case averted per 10,000 doses",
         fill = "Vaccine\nefficacy (%)" ) +
   theme_classic() +
   theme( panel.grid.major = element_line( colour="#f0f0f0" ) ) 

ggsave( "figs/vacccov_caseaverted.png", width=3.4*1.5, height=2.5*1.5, units="in" )

ggplot( resdf ) +
  geom_boxplot( aes(as.factor(100*vacc_cov), 
                     vacc, fill=as.factor(100*vacc_eff) ), 
                 position = position_dodge(width = .6) ) +
  labs( x="Vaccine coverage rate (%)", y="Number of Vaccine doses used", 
         fill = "Vaccine\nefficacy (%)" ) +
  theme_classic() +
  theme( panel.grid.major = element_line( colour="#f0f0f0" ) )
   
ggsave( "figs/vacccov_vaccdose.png", width=3.4*1.5, height=2.5*1.5, units="in" )
```

```{r}



```

```{r}
novacc <- readRDS( "out/no_vacc.RDS" )
res <- readRDS( "out/vacc_coverage.RDS" )
id <- which( res$vacc_status == 0 ) ## id to mark the sim where vaccination didn't take place because the cases were not
nvar = ncol( res$cumul_case )
ci_novacc <- rep( novacc$cumul_case, nvar )
ci <- c( novacc$cumul_case, as.vector( res$cumul_case ) )
vacc <- c( novacc$cumul_vacc, as.vector( res$cumul_vacc ) )
vacc_cov <- as.vector( sapply( c(0,30,50,70,95), function(x) rep(x,nsample)) )
dd <- tbl_df( data.frame( ids=1:length(vacc), ci =ci, vacc=vacc, vacc_cov = vacc_cov  ) )
dd$vacc_cov <- factor( dd$vacc_cov )

library( ggplot2 )
# library( ggridges )
# ggplot(dd, aes(y=as.factor(vacc_cov), x=ci)) +
#    geom_density_ridges()

ggplot( dd, aes(vacc_cov, ci) ) +
   geom_boxplot() + 
   coord_flip()+
   geom_jitter( width=0.2, alpha=0.4 ) +
   # scale_y_continuous( limits=c(0,1000) ) +
   labs( x="Vacccination coverage(%)", y="Cumulative incidence") +
   theme_classic() +
   theme( panel.grid.major = element_line( colour="#f0f0f0" ) )

# ggsave( "figs/vacc_cov_case.png", width=3.4*1.5, height=2.5*1.5, units="in" )

ggplot( dd, aes(vacc_cov, vacc) ) + 
   geom_boxplot() +
   coord_flip() +
   geom_jitter( width=0.2, alpha=0.4 ) +
   labs( x="Vacccination coverage(%)", y="Total vaccine doses") +
   theme_classic() +
   theme( panel.grid.major = element_line( colour="#f0f0f0" ) )
# ggsave( "figs/vacc_cov_vacc.png", width=3.4*1.5, height=2.5*1.5, units="in" )

```

```{r}
dd <- dd[-id,]
ve_mean <- rep( NA, nvar )
ve_med <- rep( NA, 6 )
vacc_dose_mean <- rep( NA, 6 )
vacc_dose_med <- rep( NA, 6 )
case_avt_per_dose_mean <- rep( NA, 6 )
case_avt_per_dose_med <- rep( NA, 6 )
for(i in 1: 6 ){
   ddd <- dplyr::filter( dd, 100*(i-1) < ids & ids <= 100*i )
   ci_novacc_mean <- mean( ddd$ci_novacc )

   ci_novacc_med <- median( ddd$ci_novacc )
   ci_vacc_mean <- mean( ddd$ci_vacc )
   ci_vacc_med <- median( ddd$ci_vacc )

   ve_mean[ i ] = 100*(ci_novacc_mean - ci_vacc_mean) / ci_novacc_mean
   ve_med[ i ] = 100*(ci_novacc_med - ci_vacc_med) / ci_novacc_med
   vacc_dose_mean[ i ] = mean( ddd$vacc )
   vacc_dose_med[ i ] = median( ddd$vacc )
   case_avt_per_dose_mean[ i ] <- (ci_novacc_mean - ci_vacc_mean) /vacc_dose_mean
   case_avt_per_dose_med[ i ] <- (ci_novacc_med - ci_vacc_med)/vacc_dose_med
   
}

for(i in 1: 6 ){
   ddd <- dplyr::filter( dd, 100*(i-1) < ids & ids <= 100*i ); cat(nrow(ddd),",") 
}
case_threshold <- c(rep(5,100),rep(8,100),rep(11,100),rep(14,90),rep(17,64),rep(20,34))
day_threshold <- c(rep(14,length(case_threshold)))
dd$case_threshold <- case_threshold
dd$day_threshold <- day_threshold
ddf <- dplyr::tbl_df( dd )
ddf$case_threshold <- factor( ddf$case_threshold )

library( ggplot2 )
ggplot( ddf, aes(case_threshold, ci_vacc) ) + 
   geom_boxplot( ) +
   geom_jitter( width = 0.2 ) +
   scale_y_continuous( limits=c(0,1000) ) +
   labs( x="Threshold incidence for vaccination response", y="Total cases") +
   theme_classic() +
   theme( panel.grid.major = element_line( colour="#f0f0f0" ) )

# ggsave( "figs/threshold_case_vacc.png", width=3.4*1.5, height=2.5*1.5, units="in" )

ggplot( ddf, aes(case_threshold, vacc) ) + 
   geom_boxplot( ) +
   geom_jitter( width = 0.2 ) +
   labs( x="Threshold incidence for vaccination response", y="Total vaccine doses") +
   theme_classic() +
   theme( panel.grid.major = element_line( colour="#f0f0f0" ) )
# ggsave( "figs/threshold_vacc.png", width=3.4*1.5, height=2.5*1.5, units="in" )
```
### Putting resutls into tables 
#### 
```{r}
res <- readRDS( "out/vacccov_relpop_20181219T190631.rds" )
pars <- readRDS( "out/pars_20181219T190631.rds" )

cycle <- 6# 6 comes from the vaccine coverage being 0,0,1,0.3,...0,9
index_Vacc_zero <- seq( 1, nrow(pars), by=cycle ) 
ci_control <- sapply( index_Vacc_zero, function(x) median( res[[x]]$cumul_symptom_onset ) )
ci_control <- ci_control[ rep( 1:length( ci_control ), each=cycle ) ]
# overall vaccine effectiveness
# vacc_occurred <- sapply( 1:nrow(pars), function(x) (sum(res[[x]]$vacc_day_adjusted)- pars$n_per_par[1]) )
ci <- sapply( 1:nrow(pars), function(x) median( res[[ x ]]$cumul_symptom_onset) )
vacc <- sapply( 1:nrow(pars), function(x) median( res[[ x ]]$cumul_vacc_dose ) )  
ove <- sapply( 1:nrow(pars), function(x) 100 * ( 1 - ci[ x ] / ci_control[ x ] ) )
case_averted <- sapply( 1:nrow(pars), function(x) 10000 * ( ( ci_control[ x ] - ci[ x ]  )  / vacc[ x ] ) ) #

df <- data.frame( ci=ci, vacc = vacc, ove = ove, case_averted = case_averted)

resdf <-  cbind( pars, df )

# resdf <- dplyr::filter( resdf, vacc_eff == 0.3 )
resdf <- resdf[ resdf$vacc_eff==0.7, ]

library(dplyr)

ove_summary <- resdf %>% 
  group_by( vacc_cov, vacc_scenario ) %>% 
  summarize( median = quantile( ove, probs=0.5, na.rm=TRUE ),
             lower = quantile( ove, probs=0.025, na.rm=TRUE ),
             upper = quantile( ove, probs=0.975, na.rm=TRUE ) )
vacc_summary <- resdf %>% 
  group_by( vacc_cov, vacc_scenario ) %>%  
  summarize( median = quantile( vacc, probs=0.5, na.rm=TRUE ),
             lower = quantile( vacc, probs=0.025, na.rm=TRUE ),
             upper = quantile( vacc, probs=0.975, na.rm=TRUE ) )

caseaverted_summary <- resdf %>% 
  group_by( vacc_cov, vacc_scenario ) %>% 
  summarize( median = quantile( case_averted, probs=0.5, na.rm=TRUE ),
             lower = quantile( case_averted, probs=0.025, na.rm=TRUE ),
             upper = quantile( case_averted, probs=0.975, na.rm=TRUE ) )

vacc_summary_out <- data.frame( vacc_scenario = vacc_summary$vacc_scenario,
                                case_threshold = vacc_summary$threshold_case_vacc,
                                vacc = sprintf( "%.1f (%.1f, %.1f)", 
                                                vacc_summary$median, 
                                                vacc_summary$lower, 
                                                vacc_summary$upper ) )

# write.csv( ove_summary, "out/ove_summary.csv" )
# write.csv( vacc_summary_out, "out/vacc_summary_out.csv" )
# write.csv( caseaverted_summary, "out/caseaverted_summary.csv" )

```

### Threshold case vaccine impact
```{r}
res <- readRDS( "out/thresholdcase_par1-16_20181226T171400.rds" )
pars <- readRDS( "out/pars_par1-16_20181226T171400.rds" )

cycle <- 6 # 6 comes from the vaccine coverage being 0,0,1,0.3,...0,9
index_Vacc_zero <- seq( 1, nrow(pars), by=cycle ) 
ci_control <- sapply( index_Vacc_zero, function(x) median( res[[x]]$cumul_symptom_onset ) )
ci_control <- ci_control[ rep( 1:length( ci_control ), each=cycle ) ]
dur_control <- sapply( index_Vacc_zero, function(x) median( res[[x]]$dur_outbreak ) )
dur_control <- dur_control[ rep( 1:length( dur_control ), each=cycle ) ]
num_hospital_control <- sapply( index_Vacc_zero, function(x) median( res[[x]]$num_hospital_infected ) )
num_hospital_control <- num_hospital_control[ rep( 1:length( num_hospital_control ), each=cycle ) ]

ci <- sapply( 1:nrow(pars), function(x) median( res[[ x ]]$cumul_symptom_onset) )
dur <- sapply( 1:nrow(pars), function(x) median( res[[ x ]]$dur_outbreak ) )
num_hospital <- sapply( 1:nrow(pars), function(x) median( res[[ x ]]$num_hospital_infected ) )
vacc <- sapply( 1:nrow(pars), function(x) median( res[[ x ]]$cumul_vacc_dose ) )  
ove <- sapply( 1:nrow(pars), function(x) 100 * ( 1 - ci[ x ] / ci_control[ x ] ) )
ove_dur <- sapply( 1:nrow(pars), function(x) 100 * ( 1 - dur[ x ] / dur_control[ x ] ) )

case_averted <- sapply( 1:nrow(pars), function(x) 10000 * ( ( ci_control[ x ] - ci[ x ]  )  / vacc[ x ] ) ) #
ove_hospital <- sapply( 1:nrow(pars), function(x) 100 * ( 1 - num_hospital[ x ] / num_hospital_control[ x ] ) )

df <- data.frame( ci=ci, vacc = vacc, ove = ove, dur = dur, ove_dur = ove_dur, 
                  case_averted = case_averted,
                  num_hospital = num_hospital, ove_hospital = ove_hospital )

resdf <- cbind( pars, df )
resdf <- resdf[ resdf$threshold_case_vacc == 1, ]
library(tidyverse)
ggplot( resdf ) +
  geom_boxplot( aes(as.factor(100*vacc_cov), ci, fill=as.factor(threshold_case_vacc) ),
                position = position_dodge(width = .6) ) +
  coord_cartesian( ylim=c(0,500) ) +
  labs( x="Vaccine coverage rate (%)", y="Cumulative incidence", 
         fill = "Threshold\ncase incidence" ) +
  theme_classic() +
  theme( panel.grid.major = element_line( colour="#f0f0f0" ) ) +
  facet_wrap(.~vacc_scenario)
# ggsave( "figs/vacccov_vacceff_inc.png", width=3.4*1.5, height=2.5*1.5, units="in" )

ggplot( resdf ) +
  geom_boxplot( aes(as.factor(100*vacc_cov), ove, fill=as.factor(threshold_case_vacc) ),
                position = position_dodge(width = .6) ) +
  coord_cartesian( ylim=c(-25,75) ) +
  labs( x="Vaccine coverage rate (%)", y="Reduction in outbreak size (%)", 
         fill = "Threshold\ncase incidence" ) +
  theme_classic() +
  theme( panel.grid.major = element_line( colour="#f0f0f0" ),
         panel.grid.minor = element_line( colour="#f0f0f0" )) +
  facet_wrap(.~vacc_scenario)

# ggsave( "figs/vacccov_outbreaksize_thresholdcase_scenario.png", width=3.4*3, height=2.5*3, units="in" )


ggplot( resdf ) +
  geom_boxplot( aes(as.factor(100*vacc_cov), ove_dur, fill=as.factor(threshold_case_vacc) ),
                position = position_dodge(width = .6) ) +
  # coord_cartesian( ylim=c(-25,75) ) +
  labs( x="Vaccine coverage rate (%)", y="Reduction in outbreak duration (%)", 
         fill = "Threshold\ncase incidence" ) +
  theme_classic() +
  theme( panel.grid.major = element_line( colour="#f0f0f0" ),
         panel.grid.minor = element_line( colour="#f0f0f0" )) +
  facet_wrap(.~vacc_scenario)

# ggsave( "figs/vacccov_dur_thresholdcase_scenario.png", width=3.4*3, height=2.5*3, units="in" )

ggplot( resdf ) +
  geom_boxplot( aes(as.factor(100*vacc_cov), ove_dur, fill=as.factor(vacc_scenario) ),
                position = position_dodge(width = .6) ) +
  labs( x="Vaccine coverage rate (%)", y="Reduction in outbreak duration (%)", 
         fill = "Vaccination\nscenario" ) +
  theme_classic() +
  theme( panel.grid.major = element_line( colour="#f0f0f0" ),
         panel.grid.minor = element_line( colour="#f0f0f0" ))

ggsave( "figs/vacccov_dur_vaccinationscenario.png", width=3.4*1.5, height=2.5*1.5, units="in" )


ggplot( resdf ) +
   geom_boxplot( aes(as.factor(100*vacc_cov), case_averted, 
                     fill=as.factor(threshold_case_vacc) ),
                 position = position_dodge(width = .6) ) + 
   coord_cartesian(ylim=c(-20,100)) +
   labs( x="Vaccine coverage rate (%)", y="Case averted per 10,000 doses",
         fill = "Threshold\ncase incidence" ) +
   theme_classic() +
   theme( panel.grid.major = element_line( colour="#f0f0f0" ),
          panel.grid.minor = element_line( colour="#f0f0f0" )) +
   facet_wrap(.~vacc_scenario)

# ggsave( "figs/vacccov_caseaverted_thresholdcase_scenario.png", width=3.4*1.5, height=2.5*1.5, units="in" )

ggplot( resdf ) +
  geom_boxplot( aes(as.factor(100*vacc_cov), 
                     vacc, fill=as.factor(threshold_case_vacc) ), 
                 position = position_dodge(width = .6) ) +
  labs( x="Vaccine coverage rate (%)", y="Number of Vaccine doses used", 
         fill = "Threshold\ncase incidence" ) +
  theme_classic() +
  theme( panel.grid.major = element_line( colour="#f0f0f0" ),
         panel.grid.minor = element_line( colour="#f0f0f0" )) +
  facet_wrap(.~vacc_scenario)
   
# ggsave( "figs/vacccov_vaccdose_thresholdcase_scenario.png", width=3.4*1.5, height=2.5*1.5, units="in" )

## 
ggplot( resdf ) +
  geom_boxplot( aes(as.factor(100*vacc_cov), 
                     num_hospital, fill=as.factor(threshold_case_vacc) ), 
                 position = position_dodge(width = .6) ) +
  coord_cartesian(ylim=c(-0,50)) +
  labs( x="Vaccine coverage rate (%)", y="Number of affected hospitals", 
         fill = "Threshold\ncase incidence" ) +
  theme_classic() +
  theme( panel.grid.major = element_line( colour="#f0f0f0" ) ) +
  facet_wrap(.~vacc_scenario)
   
# ggsave( "figs/vacccov_numhospitals_thresholdcase_scenario.png", width=3.4*1.5, height=2.5*1.5, units="in" )
```


### Tile plot of vaccine impact
```{r}
### Overall vaccine effectiveness
library(tidyverse)
vacc_radius <- c(30,40,50,60,70,80,90)
vacc_threhold_case <- c(5, 7, 9, 11, 13, 15)
df <- tbl_df( expand.grid( threshold=vacc_threhold_case, radius=vacc_radius ) )
df$VE <- NA
# call data
a <- readRDS( "out/no_vacc.RDS" )
b <- readRDS( "out/vacc_case_radius.RDS" )
ci_mean_novacc <- mean( a$cumul_case )
ci_mean_vacc <- colMeans( b$cumul_case )
dose_mean_vacc <- colMeans( b$cumul_vacc )

df$VE <- (ci_mean_novacc-ci_mean_vacc) / ci_mean_novacc * 100
df$ve_per_dose <- (ci_mean_novacc-ci_mean_vacc) / dose_mean_vacc * 10000
 
library(viridis)
ggplot( df, aes(radius, threshold ) ) +
   # expand_limits( x = 0, y = 0 ) +
   geom_tile( aes( fill = ve_per_dose), color = "white") +
   scale_fill_viridis() +
   labs( x="Vaccination radius", y="Threshold case incidence", fill="VE" ) +
   theme_classic() 

# ggsave( "figs/outbreak_prob.png", width=3.4*1.5, height=2.5*1.5, units="in" )

cut_off <- seq( 1, 30, 1 )
time <- 1:30
df <- tbl_df( expand.grid( time=time, cut_off=cut_off ) )
df$prob <- NA
ci <- tbl_df( cumul_symptom_onset ) 
index = 1
for( i in 1:length(cut_off) ){
   for( j in 1:length(time) ) {
      con_1 <- function(x) x[ j ] > cut_off[ i ]
      con_2 <- function(x) tail(x,1) > 200
      d1 <- ci %>% select_if( con_1 )
      d2 <- d1 %>% select_if( con_2 )
      if( ncol(d1) > 0 ){ 
         df[ index, "prob" ] <- ncol( d2 ) / ncol( d1 )
      }
      index <- index + 1
   }
}

library(viridis)
ggplot( df, aes(time, cut_off ) ) +
   expand_limits( x = 0, y = 0 ) +
   geom_tile( aes( fill = prob), color = "white") +
   scale_fill_viridis() +
   labs( x="Day", y="Cut-off", fill="Prob" ) +
   theme_classic() 

ggsave( "figs/outbreak_prob.png", width=3.4*1.5, height=2.5*1.5, units="in" )

```
```{r}
ci <- tbl_df( cumul_symptom_onset[ nrow(cumul_symptom_onset), ] )
ggplot( ci ) +
   geom_histogram( aes( x=value ), binwidth = 10 ) +
   labs( x="Outbreak size", y="Frequency" ) +
   theme_classic() +
   theme( panel.grid.major = element_line( colour="grey94" ) )

# ggsave( "figs/outbreak_size_dist_ts0p1.png", width=3.4*1.5, height=2.5*1.5, units="in" )

```

#### Outbreak size distribution upon importation of a case
```{r}
res <- readRDS( "out/importation_20190129T103006.rds" )
ci <- sapply( seq_along(res), function(x) unlist( res[[x]]$cumul_symptom_onset ) ) 
d <- data.frame( ci = as.vector( ci ) )
library( ggplot2 )
ggplot( d ) +
   geom_histogram( aes( x=ci ), binwidth = 50 ) +
   labs( x="Outbreak size", y="Frequency" ) +
   theme_classic() +
   theme( panel.grid.major = element_line( colour="grey94" ) )
#    
# ggsave( "figs/outbreak_size_distribution.png", width=3.4*1.5, height=2.5*1.5, units="in" )

ggplot( d, aes(ci) ) +
  stat_ecdf( geom = "point" ) +
  stat_ecdf( geom = "step" ) +
  geom_vline( xintercept = 187 ) +
  geom_hline( yintercept = ecdf( d$ci )( 187 ) ) +
  geom_point( aes(187, ecdf( d$ci )( 187 )), color="red", size=1.5, inherit.aes=FALSE ) +
  labs( x="Outbreak size", y="Cumulative probabability" ) +
  theme_classic() +
  theme( panel.grid.major = element_line( colour="grey94" ) ) + 
  annotate( "text", x = 10000, y = ecdf( d$ci )( 150 ), 
            label = paste0( "2015 Korean MERS-CoV outbreak (", sprintf("%.1f",ecdf( d$ci )( 187 )), " percentile)"), size=3 )
# ggsave( "figs/outbreak_size_distribution_cumul.png", width=3.4*1.5, height=2.5*1.5, units="in" )  

ci_ordered <- d$ci[ order(d$ci) ]
cdf <- ecdf( d$ci )
cdf_val <- cdf( ci_ordered )
d <- bind_cols( data_frame( t=1:nrow(daily_symptom_onset) ), tbl_df(daily_symptom_onset) ) 
d_long <- gather( d, key, val, -t)
ggplot( d_long ) +
   geom_line( aes( x=t, y=val, color=key ), alpha=0.2, show.legend=FALSE ) + 
   scale_color_manual( values=rep("black",1000) ) +
   labs( x="Day", y="Symptom onset" ) +
   theme_classic() +
   theme( panel.grid.major = element_line( colour="grey94" ) )

# ggsave( "figs/outbreak_size_distribution.png", width=3.4*1.5, height=2.5*1.5, units="in" )
```
#### Statistics on importation-driven outbreaks
```{r}
res <- readRDS( "out/importation_20181221T143502.rds" )

ci <- sapply( seq_along(res), function(x) unlist( res[[x]]$cumul_symptom_onset ) ) 
d <- data.frame( ci = as.vector( ci ) )
table( d$ci )
  
```
